---
title: "Primary election"
author: "Victor Kostyuk"
date: "August 20, 2016"
output: html_document
---

```{r setup, echo = TRUE, message = FALSE}
# Load libraries
library(dplyr)
library(tidyr)
library(randomForest)
library(forestFloor)
knitr::opts_chunk$set(fig.path='Figs/', echo = TRUE, warning = FALSE, 
                      message = FALSE, results = "hide")
```

```{r load_data, cache = TRUE}
results <- read.csv("csv/primary_results.csv", header = TRUE)
county <- read.csv("csv/county_facts.csv", header = TRUE)
```

```{r clean_and_combine, cache = TRUE}
# Widen results, one county per row
df <- results %>%
  select(-party) %>%
  gather(variable, value, votes:fraction_votes) %>% 
  unite(tmp, candidate, variable) %>% 
  spread(tmp, value) %>% 
  # Join county features
  inner_join(county, by = "fips")
```

# Visualizing partial response and feature interaction in a random forest

We define a response `Clinton_over_trump` as the difference of Clinton's fraction of votes and Trump's.

Then, we use a random forest to model this response and explore variable importance and interaction using the `forestFloor` package.

```{r clint_over_trump, cache = TRUE}
df$Clinton_over_trump <- 
  df$`Hillary Clinton_fraction_votes` - df$`Donald Trump_fraction_votes`

df_no_na <- df[!(is.na(df$Clinton_over_trump)), ]

x <- df_no_na[, 39:89]
y <- df_no_na$Clinton_over_trump

rf_ct <- randomForest(x, y, scale = FALSE, ntree = 1000,
                   keep.forest = TRUE, keep.inbag = TRUE,
                   importance = TRUE)

# Increase in node purity importance
varImpPlot(rf_ct, type = 2)
```

```{r, cache = TRUE}
# Forest floor -- can't figure out how to use increse in MSE importance,
# so using increase in node purity (somewhat worse)
ff <- forestFloor(rf_ct, x, calc_np = FALSE)
```

Color gradient corresponds to the first most important feature:

```{r}
Col <- fcol(ff, 1, orderByImportance = TRUE)
plot(ff, col = Col, plot_seq = 1:9, 
     orderByImportance = TRUE)
```

Color gradient corresponds to the second most important feature:

```{r}
Col <- fcol(ff, 2, orderByImportance = TRUE)
plot(ff, col = Col, plot_seq = 1:9, 
     orderByImportance = TRUE)
```

Color gradient corresponds to first principal component of the first three most important features:

```{r}
Col <- fcol(ff, 1:3, max.df = 1, orderByImportance = TRUE)
plot(ff, col = Col, plot_seq = 1:9, 
     orderByImportance = TRUE)
```

Color gradient corresponds to incompe per capita:

```{r}
Col <- fcol(ff, which(colnames(x) == "Income_per_capita"), orderByImportance = FALSE)
plot(ff, col = Col, plot_seq = 1:9, 
     orderByImportance = TRUE)
```

Here's how to see an rgl 3d plot across two features.

```{r, eval = FALSE}
# The x-y plane is spanned by the first and ninth most important feature,
# color grandient corresponds to first feature
show3d(ff,c(1,9), col = fcol(ff, 1, orderByImportance = TRUE), 
       orderByImportance = TRUE)
```